---
title: "COVID-19 - Forecasting and compare machine learning models"
author: "Adrien BORDERON"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Load the following libraries. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(modeltime)
library(timetk)   
library(lubridate)
library(readr)
library(dplyr)
library(tidyr)
library(glmnet)
library(randomForest)
```

# Data on COVID-19 (coronavirus) by _Our World in Data_

COVID-19 dataset is a collection of the COVID-19 data maintained by [_Our World in Data_](https://ourworldindata.org/coronavirus). It is updated daily and includes data on confirmed cases, deaths, and testing, as well as other variables of potential interest.

We’ll start with a covid daily time series data set that includes **new daily cases**. We’ll simplify the data set to a univariate time series with columns, “date” and “value”.

```{r warning=FALSE}
data_daily  <- read.csv("~/Workdir/COVID-19-models-forecasting/data/owid-covid-data.csv")
data_daily <- data_daily %>% select(date, new_cases) %>% drop_na(new_cases)

time_series_covid19_confirmed_global <- aggregate(data_daily$new_cases, by=list(date=data_daily$date), FUN=sum)

covid19_confirmed_global_tbl <- time_series_covid19_confirmed_global %>%
  select(date, x) %>%
  rename(value=x) 

covid19_confirmed_global_tbl$date <- as.Date(covid19_confirmed_global_tbl$date, format = "%Y-%m-%d")
covid19_confirmed_global_tbl <- covid19_confirmed_global_tbl[covid19_confirmed_global_tbl[["date"]] >= "2020-01-17", ]

covid19_confirmed_global_tbl <- as.tibble(covid19_confirmed_global_tbl)
covid19_confirmed_global_tbl
```

## Show COVID-19 new daily cases

```{r, fig.width=14, fig.height=8}
covid19_confirmed_global_tbl %>%
  plot_time_series(date, value, .interactive = FALSE)
```

# Train / Test

Split time series into training and testing sets. We use the last ****5 days** of data as the testing set.

## Train/test split time serie

```{r, fig.width=14, fig.height=8}
splits <- covid19_confirmed_global_tbl %>%
  time_series_split(assess = "5 days", cumulative = TRUE)

splits %>%
  tk_time_series_cv_plan() %>%
  plot_time_series_cv_plan(date, value, .interactive = FALSE)
```

# Modeling

## Automatic Models

Several models to see this process :

### Auto ARIMA

Auto Arima Model fitting process.

```{r}
model_fit_arima <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(value ~ date, training(splits))

model_fit_arima
```

### Prophet

Prophet Model fitting process.

```{r}
model_fit_prophet <- prophet_reg() %>%
  set_engine("prophet", yearly.seasonality = TRUE) %>%
  fit(value ~ date, training(splits))

model_fit_prophet
```

## Machine Learning Models

### Preprocessing Recipe

```{r}
recipe_spec <- recipe(value ~ date, training(splits)) %>%
  step_timeseries_signature(date) %>%
  step_rm(contains("am.pm"), contains("hour"), contains("minute"),
          contains("second"), contains("xts")) %>%
  step_fourier(date, period = 365, K = 5) %>%
  step_dummy(all_nominal())

recipe_spec %>% prep() %>% juice()
```

With a recipe in-hand, we can set up our machine learning pipelines.

### Elastic Net

Making an Elastic NET model

```{r}
model_spec_glmnet <- linear_reg(penalty = 0.01, mixture = 0.5) %>%
  set_engine("glmnet")
```

Fitted workflow

```{r}
workflow_fit_glmnet <- workflow() %>%
  add_model(model_spec_glmnet) %>%
  add_recipe(recipe_spec %>% step_rm(date)) %>%
  fit(training(splits))

```

### Random Forest

Fit a Random Forest

```{r}
model_spec_rf <- rand_forest(trees = 500, min_n = 50) %>%
  set_engine("randomForest")

workflow_fit_rf <- workflow() %>%
  add_model(model_spec_rf) %>%
  add_recipe(recipe_spec %>% step_rm(date)) %>%
  fit(training(splits))
```

### Prophet Boost

Set the model up using a workflow

```{r warning=FALSE}
model_spec_prophet_boost <- prophet_boost() %>%
  set_engine("prophet_xgboost", yearly.seasonality = TRUE) 

workflow_fit_prophet_boost <- workflow() %>%
  add_model(model_spec_prophet_boost) %>%
  add_recipe(recipe_spec) %>%
  fit(training(splits))

workflow_fit_prophet_boost
```

# Model evaluation and selection

Organize the models with IDs and creates generic descriptions.

```{r}
model_table <- modeltime_table(
  model_fit_arima, 
  model_fit_prophet,
  workflow_fit_glmnet,
  workflow_fit_rf,
  workflow_fit_prophet_boost
) 

model_table
```

## Calibration

Quantify error and estimate confidence intervals.

```{r}
calibration_table <- model_table %>%
  modeltime_calibrate(testing(splits))

calibration_table
```

# Forecast (Testing Set)

Visualize the testing predictions (forecast).

```{r, fig.width=14, fig.height=8}
calibration_table %>%
  modeltime_forecast(actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE)
```

## Accuracy (Testing Set)

Calculate the testing accuracy to compare the models.

```{r}
calibration_table %>%
  modeltime_accuracy() %>%
  table_modeltime_accuracy(.interactive = FALSE)
```

## Analyze Results

From the accuracy measures and forecast results, we see that:

* RANDOMFOREST model is not a good fit for this data.
* The best model are GLMNET and Prophet + XGBoost
Let’s exclude the RANDOMFOREST from our final model, then make future forecasts with the remaining models.


# Refit and Forecast Forward (Global in world)

```{r, fig.width=14, fig.height=8}
calibration_table %>%
  # Remove RANDOMFOREST model with low accuracy
  filter(.model_id != 4) %>%
  
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "Compare all models forecast Plot", .y_lab = "New daily cases")
```

## ARIMA model forecast

```{r, fig.width=14, fig.height=8}
calibration_table %>%
  filter(.model_id == 1) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "ARIMA model forecast Plot", .y_lab = "New daily cases")
```

## PROPHET model forecast

```{r, fig.width=14, fig.height=8}
calibration_table %>%
  filter(.model_id == 2) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "PROPHET model forecast Plot", .y_lab = "New daily cases")
```

## GLMNET model forecast

```{r, fig.width=14, fig.height=8}
calibration_table %>%
  filter(.model_id == 3) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "GLMNET model forecast Plot", .y_lab = "New daily cases")
```

## PROPHET + XGBOOST model forecast

```{r fig.height=8, fig.width=14, warning=FALSE}
calibration_table %>%
  filter(.model_id == 5) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "PROPHET + XGBOOST model forecast Plot", .y_lab = "New daily cases")
```


\newpage

# Data on COVID-19 (coronavirus) only in FRANCE

Filter COVID-19 dataset in **France**.

```{r echo=FALSE}
data_daily  <- read.csv("~/Workdir/COVID-19-models-forecasting/data/owid-covid-data.csv")
data_daily <- data_daily[data_daily$location == 'France',]
data_daily <- data_daily %>% select(date, new_cases) %>% drop_na(new_cases)

time_series_covid19_confirmed_global <- aggregate(data_daily$new_cases, by=list(date=data_daily$date), FUN=sum)

covid19_confirmed_global_tbl <- time_series_covid19_confirmed_global %>%
  select(date, x) %>%
  rename(value=x) 

covid19_confirmed_global_tbl$date <- as.Date(covid19_confirmed_global_tbl$date, format = "%Y-%m-%d")
covid19_confirmed_global_tbl <- covid19_confirmed_global_tbl[covid19_confirmed_global_tbl[["date"]] >= "2020-02-26", ]

covid19_confirmed_global_tbl <- as.tibble(covid19_confirmed_global_tbl)
covid19_confirmed_global_tbl
```

## Show FRANCE COVID-19 new daily cases

```{r echo=FALSE, fig.height=8, fig.width=14}
covid19_confirmed_global_tbl %>%
  plot_time_series(date, value, .interactive = FALSE)
```

# Train / Test

## Train/test split time serie

```{r echo=FALSE, fig.height=8, fig.width=14}
splits <- covid19_confirmed_global_tbl %>%
  time_series_split(assess = "5 days", cumulative = TRUE)

splits %>%
  tk_time_series_cv_plan() %>%
  plot_time_series_cv_plan(date, value, .interactive = FALSE)
```

# Modeling

## Automatic Models

### Auto ARIMA

Auto Arima Model fitting process.

```{r}
model_fit_arima <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(value ~ date, training(splits))

model_fit_arima
```

### Prophet

Prophet Model fitting process.

```{r}
model_fit_prophet <- prophet_reg() %>%
  set_engine("prophet", yearly.seasonality = TRUE) %>%
  fit(value ~ date, training(splits))

model_fit_prophet
```

## Machine Learning Models

### Preprocessing Recipe

```{r echo=FALSE}
recipe_spec <- recipe(value ~ date, training(splits)) %>%
  step_timeseries_signature(date) %>%
  step_rm(contains("am.pm"), contains("hour"), contains("minute"),
          contains("second"), contains("xts")) %>%
  step_fourier(date, period = 365, K = 5) %>%
  step_dummy(all_nominal())

recipe_spec %>% prep() %>% juice()
```

### Elastic Net

Making an Elastic NET model

```{r}
model_spec_glmnet <- linear_reg(penalty = 0.01, mixture = 0.5) %>%
  set_engine("glmnet")
```

Fitted workflow

```{r}
workflow_fit_glmnet <- workflow() %>%
  add_model(model_spec_glmnet) %>%
  add_recipe(recipe_spec %>% step_rm(date)) %>%
  fit(training(splits))

```

### Random Forest

Fit a Random Forest

```{r echo=FALSE}
model_spec_rf <- rand_forest(trees = 500, min_n = 50) %>%
  set_engine("randomForest")

workflow_fit_rf <- workflow() %>%
  add_model(model_spec_rf) %>%
  add_recipe(recipe_spec %>% step_rm(date)) %>%
  fit(training(splits))
```

### Prophet Boost

Set the model up using a workflow

```{r echo=FALSE}
model_spec_prophet_boost <- prophet_boost() %>%
  set_engine("prophet_xgboost", yearly.seasonality = TRUE) 

workflow_fit_prophet_boost <- workflow() %>%
  add_model(model_spec_prophet_boost) %>%
  add_recipe(recipe_spec) %>%
  fit(training(splits))

workflow_fit_prophet_boost
```

# Model evaluation and selection

Organize the models with IDs and creates generic descriptions.

```{r echo=FALSE}
model_table <- modeltime_table(
  model_fit_arima, 
  model_fit_prophet,
  workflow_fit_glmnet,
  workflow_fit_rf,
  workflow_fit_prophet_boost
) 

model_table
```

## Calibration

Quantify error and estimate confidence intervals.

```{r echo=FALSE}
calibration_table <- model_table %>%
  modeltime_calibrate(testing(splits))

calibration_table
```

# FRANCE Forecast (Testing Set)

Visualize the testing predictions (forecast).

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  modeltime_forecast(actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE)
```

## Accuracy (Testing Set)

Calculate the testing accuracy to compare the models.

```{r echo=FALSE}
calibration_table %>%
  modeltime_accuracy() %>%
  table_modeltime_accuracy(.interactive = FALSE)
```

## Analyze Results

From the accuracy measures and forecast results, we see that:

* RANDOMFOREST model is not a good fit for this data.
* The best model is PROPHET W/ XGBOOST ERRORS
Let’s exclude the RANDOMFOREST from our final model, then make future forecasts with the remaining models.


# Refit and Forecast Forward (FRANCE ONLY)

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  # Remove RANDOMFOREST model with low accuracy
  filter(.model_id != 4) %>%
  
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "Compare all models forecast Plot", .y_lab = "FRANCE New daily cases")
```

## ARIMA model forecast

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  filter(.model_id == 1) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "ARIMA model forecast Plot", .y_lab = "FRANCE New daily cases")
```

## PROPHET model forecast

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  filter(.model_id == 2) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "PROPHET model forecast Plot", .y_lab = "FRANCE New daily cases")
```

## GLMNET model forecast

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  filter(.model_id == 3) %>%
  
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "GLMNET model forecast Plot", .y_lab = "FRANCE New daily cases")
```

## PROPHET + XGBOOST model forecast

```{r echo=FALSE, fig.height=8, fig.width=14}
calibration_table %>%
  filter(.model_id == 5) %>%
  # Refit and Forecast Forward
  modeltime_refit(covid19_confirmed_global_tbl) %>%
  modeltime_forecast(h = "3 months", actual_data = covid19_confirmed_global_tbl) %>%
  plot_modeltime_forecast(.interactive = FALSE, .title = "PROPHET + XGBOOST model forecast Plot", .y_lab = "FRANCE New daily cases")
```
